---
title: "Extended analysis on FLARE WP2 Experiment 1"
format:
  html:
    code-fold: true
editor: visual
---

## Load data

We keep only the data from participants that finished the experiment (n = 77)

```{r message=FALSE, warning=FALSE}
# Load necessary libraries
library(tidyverse)
library(afex)
library(kableExtra)

# Define paths
exp_code <- "exp_3"
data_dir <- file.path("..", "data", exp_code, "derivatives")

# Load and combine all .csv files with exactly 195 rows
file_list <- list.files(data_dir, pattern = "\\.(csv)$", full.names = TRUE)

# Read as character (to avoid issues when binding) and keep only those with 195 rows
dfs <- file_list  %>% 
  map(~ read_csv(.x, col_types = cols(.default = "c")))  %>% 
  keep(~ nrow(.) == 195)

data <- bind_rows(dfs) %>% 
  mutate(
    subject = as.factor(subject),
    rt = as.numeric(rt),
    dot_acc = as.numeric(dot_acc),
    verbal_acc_corrected = as.numeric(verbal_acc_corrected),
    semantic_distance = as.numeric(semantic_distance),
    manipulation = as.factor(manipulation),
    phase = as.factor(phase),
    image_type = as.factor(image_type),
    block_type = as.factor(block_type),
    phase = fct_relevel(phase, "pre", "post")

  ) %>% 
  select(subject,rt,dot_acc,verbal_acc_corrected,semantic_distance,manipulation,
         phase,image_type,block_type,image_id)



head(data)
```

## Descriptive plots

```{r}
data %>%
  filter(rt > 100, rt < 5000, dot_acc == 1) %>% 
  ggplot(aes(x = rt, fill = manipulation)) +
  geom_density(alpha = 0.3) +
  labs(
    title = "RT Distribution in the dot task",
    x = "Reaction Time (ms)",
    y = "Density"
  ) +
  facet_wrap(vars(image_type)) +
  theme_minimal()

data %>%
  filter(rt > 100, rt < 5000) %>% 
  ggplot(aes(x = dot_acc, fill = manipulation)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Accuracy Distribution in the dot task",
    x = "accuracy",
    y = "Density"
  ) +
  facet_wrap(vars(image_type)) +
  theme_minimal()

```

Average number of trials per participant and condition (manipulation x phase x image_type) for the RT analysis:

```{r}
trial_counts <- data %>%
  filter(
    block_type == "dot",
    dot_acc == 1,
    rt > 100, rt < 4000
  ) %>%
  count(subject, manipulation, phase, image_type, name = "n_trials")

trial_counts <- trial_counts %>%
  group_by(manipulation, phase, image_type) %>%
  summarise(
    mean_trials = mean(n_trials),
    sd_trials = sd(n_trials),
    .groups = "drop"
  ) %>%
  arrange(manipulation, phase, image_type)
kbl(trial_counts)
```

It seems like we are running short of observations for catch trials, so it might be better to omit them. Doing so for now...

## Dot task

### Reaction times

```{r warning=FALSE}
data_dot_rt <- data %>% 
  filter(image_type == "target",
         block_type == "dot",
        dot_acc == 1,
        rt > 100, rt < 4000
         )

plot_data <- data_dot_rt %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_rt = mean(rt), .groups = "drop")

# Plot
ggplot(plot_data, aes(x = phase, y = mean_rt)) +
  geom_jitter(aes(color = manipulation), width = 0.1, alpha = 0.4, size = 1.5) +  # individual dots
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "line", size = 1.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun.data = mean_se,
               geom = "errorbar", width = 0.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "point", size = 3) +
  facet_wrap(~ manipulation) +
  labs(
    title = "RT by Phase, Faceted by Manipulation",
    x = "Phase",
    y = "Mean RT (ms)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

If we do some stats, there's only a main effect of Phase

```{r message= FALSE}


# Summarise mean RT per subject × phase × manipulation
aov_data <- data_dot_rt %>%
  filter(rt > 100, rt < 4000) %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_rt = mean(rt), .groups = "drop") %>%
  mutate(phase = factor(phase, levels = c("pre", "post")))

anova_rt <- aov_ez(
  id = "subject",
  dv = "mean_rt",
  within = c("phase", "manipulation"),
  data = aov_data,
  type = 3,
  anova_table = list(correction = "GG", es = "pes")
)

kbl(anova_rt$anova_table, caption = "ANOVA on RTs", booktabs = T, escape = F, digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```

Now do it for images that were **recognized** during the disambiguation phase

```{r}
# Get recognized image trials during the clear (color) phase
recognized_images <- data %>%
  filter(
    block_type != "dot",
    phase == "color",
    image_type == "target",
    semantic_distance < 0.25
  ) %>%
  select(subject, manipulation, image = image_id) %>%  
  distinct()

data_dot_rt_recognized <- data_dot_rt %>%
  filter(image_type == "target") %>%  # just target images
  inner_join(recognized_images, by = c("subject", "manipulation", "image_id" = "image"))

plot_data <- data_dot_rt_recognized %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_rt = mean(rt), .groups = "drop")

# Plot
ggplot(plot_data, aes(x = phase, y = mean_rt)) +
  geom_jitter(aes(color = manipulation), width = 0.1, alpha = 0.4, size = 1.5) +  # individual dots
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "line", size = 1.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun.data = mean_se,
               geom = "errorbar", width = 0.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "point", size = 3) +
  facet_wrap(~ manipulation) +
  labs(
    title = "RT by Phase, Faceted by Manipulation",
    x = "Phase",
    y = "Mean RT (ms)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Summarise mean RT per subject × phase × manipulation
aov_data <- data_dot_rt_recognized %>%
  filter(rt > 100, rt < 4000) %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_rt = mean(rt), .groups = "drop") %>%
  mutate(phase = factor(phase, levels = c("pre", "post")))

anova_rt <- aov_ez(
  id = "subject",
  dv = "mean_rt",
  within = c("phase", "manipulation"),
  data = aov_data,
  type = 3,
  anova_table = list(correction = "GG", es = "pes")
)

kbl(anova_rt$anova_table, caption = "ANOVA on RTs", booktabs = T, escape = F, digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria") 
```

### Accuracy scores

```{r warning=FALSE}
data_dot_acc <- data %>% 
  filter(image_type == "target",
         block_type == "dot",
         )

plot_data <- data_dot_acc %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_acc = mean(dot_acc), .groups = "drop")

# Plot
ggplot(plot_data, aes(x = phase, y = mean_acc)) +
  geom_jitter(aes(color = manipulation), width = 0.1, alpha = 0.4, size = 1.5) +  # individual dots
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "line", size = 1.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun.data = mean_se,
               geom = "errorbar", width = 0.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "point", size = 3) +
  facet_wrap(~ manipulation) +
  labs(
    title = "Accuracy by Phase, Faceted by Manipulation",
    x = "Phase",
    y = "Mean accuracy"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
If we do some stats, there's only a main effect of Phase

```{r message= FALSE}


# Summarise mean RT per subject × phase × manipulation
aov_data <- data_dot_acc %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_acc = mean(dot_acc), .groups = "drop") %>%
  mutate(phase = factor(phase, levels = c("pre", "post")))

anova_acc <- aov_ez(
  id = "subject",
  dv = "mean_acc",
  within = c("phase", "manipulation"),
  data = aov_data,
  type = 3,
  anova_table = list(correction = "GG", es = "pes")
)

kbl(anova_acc$anova_table, caption = "ANOVA on Accuracy", booktabs = T, escape = F, digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```

Now do it for images that were **recognized** during the disambiguation phase

```{r}
# Get recognized image trials during the clear (color) phase
recognized_images <- data %>%
  filter(
    block_type != "dot",
    phase == "color",
    image_type == "target",
    semantic_distance < 0.25
  ) %>%
  select(subject, manipulation, image = image_id) %>%  
  distinct()

data_dot_acc_recognized <- data_dot_acc %>%
  filter(image_type == "target") %>%  # just target images
  inner_join(recognized_images, by = c("subject", "manipulation", "image_id" = "image"))

plot_data <- data_dot_acc_recognized %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_acc = mean(dot_acc), .groups = "drop")

# Plot
ggplot(plot_data, aes(x = phase, y = mean_acc)) +
  geom_jitter(aes(color = manipulation), width = 0.1, alpha = 0.4, size = 1.5) +  # individual dots
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "line", size = 1.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun.data = mean_se,
               geom = "errorbar", width = 0.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean,
               geom = "point", size = 3) +
  facet_wrap(~ manipulation) +
  labs(
    title = "Accuracy by Phase, Faceted by Manipulation",
    x = "Phase",
    y = "Accuracy"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Summarise mean RT per subject × phase × manipulation
aov_data <- data_dot_acc_recognized %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_acc = mean(dot_acc), .groups = "drop") %>%
  mutate(phase = factor(phase, levels = c("pre", "post")))

anova_rt <- aov_ez(
  id = "subject",
  dv = "mean_acc",
  within = c("phase", "manipulation"),
  data = aov_data,
  type = 3,
  anova_table = list(correction = "GG", es = "pes")
)


kbl(anova_rt$anova_table, caption = "ANOVA on Accuracy", booktabs = T, escape = F, digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria") 
```


## Naming task

### Verbal accuracy
```{r}
data_recog <- data %>%
  filter(block_type != "dot", image_type == "target", phase != 'color')
data_recog %>%
  group_by(subject, phase, manipulation) %>%
  summarise(acc = mean(verbal_acc_corrected, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = phase, y = acc)) +
  geom_jitter(aes(color = manipulation), width = 0.1, alpha = 0.3) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean, geom = "point", size = 3) +
  stat_summary(aes(group = manipulation, color = manipulation), fun.data = mean_se, geom = "errorbar", width = 0.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean, geom = "line") +
  facet_wrap(~ manipulation) +
  labs(
    title = "Verbal Identification Accuracy by Phase",
    y = "Accuracy (Proportion Correct)",
    x = "Phase"
  ) +
  theme_minimal() +
  theme(legend.position = "none")



aov_data <- data_recog %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_acc = mean(verbal_acc_corrected), .groups = "drop") %>%
  mutate(phase = factor(phase, levels = c("pre", "post")))

library(rstatix)
res.aov <- anova_test(                                 
  data = aov_data,                                 
  dv = mean_acc,                                             
  wid = subject,                                       
  within = c(phase, manipulation),
  effect.size = "pes")


kbl(res.aov$ANOVA, caption = "ANOVA on naming verbal accuracy", booktabs = T, escape = F, digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```

### Semantic distance
```{r}
data_recog <- data %>%
  filter(block_type != "dot", image_type == "target", phase != 'color')
data_recog %>%
  group_by(subject, phase, manipulation) %>%
  summarise(acc = mean(semantic_distance, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = phase, y = acc)) +
  geom_jitter(aes(color = manipulation), width = 0.1, alpha = 0.3) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean, geom = "point", size = 3) +
  stat_summary(aes(group = manipulation, color = manipulation), fun.data = mean_se, geom = "errorbar", width = 0.2) +
  stat_summary(aes(group = manipulation, color = manipulation), fun = mean, geom = "line") +
  facet_wrap(~ manipulation) +
  labs(
    title = "Semantic distance by Phase",
    y = "Semantic distance",
    x = "Phase"
  ) +
  theme_minimal() +
  theme(legend.position = "none")



aov_data <- data_recog %>%
  group_by(subject, phase, manipulation) %>%
  summarise(mean_acc = mean(semantic_distance, na.rm = TRUE), .groups = "drop") %>%
  mutate(phase = factor(phase, levels = c("pre", "post")))

res.aov <- anova_test(                                 
  data = aov_data,                                 
  dv = mean_acc,                                             
  wid = subject,                                       
  within = c(phase, manipulation),
  effect.size = "pes")


kbl(res.aov$ANOVA, caption = "ANOVA on naming semantic_distance", booktabs = T, escape = F, digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```
